Gravitational field calculation in spherical coordinates using variable densities in depth

### 2018-04-17 manuscript/manuscript.tex
Commented out old text in manuscript

\subsubsection{Linear Density}

If the density of the spherical shell is a linear function on the radial coordinate, i.e.:

\begin{equation}
    \rho(r') = ar' + b,
\end{equation}

\noindent then the gravity potential generated on any computation point at distance $r$ from its centre can be obtained from the following expression:

The first term on equation \ref{eq:shell-pot-linear} reproduces the potential generated by a spherical shell with variable density $\rho(r') = ar'$, while the second one constitutes the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
So the potential can be read as the one generated by a combination of both shells.

\subsubsection{Exponential Density}

If, on the other hand, the spherical shell has an exponential density in the radial coordinate,

\noindent the potential (calculated through SymPy \citep{sympy2017}) can be written as follows:

@@

\section{Determination of distance-size ratio}

In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy, we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
The first one constitutes the simplest variation, the second one represents a higher varying function, while the later one tries to emulate a contact between two different media (e.g. crust and mantle interface).

We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
More information about this grids can be found on Table \ref{tab:grids}.

We computed these gravity fields on every point of the four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5, without applying the density-based discretization algorithm in order to isolate the behaviour of the adaptive discretization one.
Then we calculated the maximum difference between these results and the value that the analytical solution assumes at the same height.

Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude.
That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.

Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.

\subsection{Linear Density}
Firstly,  we proposed a spherical shell with linear density in such way it assumes 2670~kg/m$^3$ on the surface of the Earth and 3300~kg/m$^3$ at the inner radius:

Then we computed the difference between the numerical model with its corresponding analytical solution (equation \ref{eq:density-linear}) for a thick and a thin spherical shell of 1km and 35km of thickness, respectively.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/linear-D-thin.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}. The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-linear-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/linear-D-thick.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}. The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-linear-thick}
\end{figure}

To test the model in case of an exponential density we proposed the following function:

\noindent and the value of $b$ equal to the thickness of the spherical shell we will use to perform the test.
In this way, the density function assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and inner radii, respectively.
The analytical solution of the gravity potential is easily obtained as the sum of the ones for a pure exponential function (like the one in equation \ref{eq:density-exp}) and for an homogeneous spherical shell.

We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, without applying the density-based discretization algorithm.
On both cases, the constant $b$ is equal to the thickness of the shell: 1km for the thin shell and 35km for the thick one.
The results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
Once again, we exclude the ones corresponding to $g_{xx}$ and $g_{yy}$ due to their similarity with the ones obtained for $g_{zz}$, although they can be found in the repository.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/exponential-D-thin.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}. The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-exp-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/exponential-D-thick.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}. The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-exp-thick}
\end{figure}

In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.

Moreover, we can see that the differences for the potential and the $g_z$ component of the gradient keep constant even for high values of $D$.
Our interpretation is that the modified adaptive discretization algorithm is limited when increasing the accuracy for this density function, supporting the need of a density-based discretization algorithm.

Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$ when the density-based discretization algorithm is not applied.
In order to do this we proposed a thin spherical shell (1km of thickness) with a set of density functions, each one with a different value of $b$.
We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the corresponding analytical solution.
These differences were calculated for each proposed density function and for the same set of values of $D$ we have been using on the previous tests, obtaining one curve for each value of $b$.

On Figure \ref{fig:D-exp-power-thin}a we can see the density functions used to perform these tests and its corresponding results on \ref{fig:D-exp-power-thin}b .
Taking into account that lower values of $b$ implies a more varying function, we can see that for high varying functions the numerical model does not perform well: for $b=100$m the difference is almost 100\%.

We can conclude that even when the density-based discretization algorithm is not applied, the numerical approximation performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, i.e. for low density variations.
On the other hand, a density-based discretization algorithm is needed in order to guarantee the accuracy of the approximation in case of a high varying density function.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/exponential-b-thin.pdf}
\caption{
    Results of the distance-size ratio determination for different exponential density functions in case of a spherical shell of 1km of thickness.
    (a) Density functions used in the tests, each one with a different value of $b$.
    (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function. Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy.
}
\label{fig:D-exp-power-thin}
\end{figure}

\subsection{Discontinuous Density}

As a final case, we performed the same test with a discontinuous density function:

\begin{equation}
    \rho(r') = 
    \begin{cases}
        \rho_A & \text{if} \, R_1 \leq r' < R_c \\
        \rho_B & \text{if} \, R_c \leq r' \leq R_2 \\
    \end{cases}
\label{eq:density-discontinuous}
\end{equation}

\noindent where $\rho_A$ and $\rho_B$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the radius on which the density change takes place.

The analytical solution for this kind of density is a linear combination of two homogeneous shells with densities $\rho_A$ and $\rho_B$:

\begin{equation}
    V(r) = \frac{4}{3} \frac{\pi G}{r}
    \left[ \rho_A (R_c^3 - R_1^3) +
    \rho_B (R_2^3 - R_c^3) \right]
\end{equation}

Lets assume that $R_c = (R_1 + R_2)/3$ and perform the test for a thin shell of 1km of thickness without applying the density-based discretization algorithm.
Its results can be seen in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what leads us to conclude that a density-based discretization algorithm is needed in order to guarantee the accuracy of the model.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/discontinuous-D-asym.pdf}
\caption{
    Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
}
\label{fig:D-discont-asym}
\end{figure}

@@

\section{Determination of delta ratio}

In the light of the previous results, a density-based discretization algorithm is needed in order to achieve an acceptable accuracy in those cases on which the modified adaptive discretization algorithm is not sufficient to take into account the numerical error due to density variations.

Besides the implementation of this new algorithm, we must determine the maximum value of the delta ratio $\delta$ that produces an accuracy bellow a predefined threshold in order to guarantee the good performance of the numerical model.

To do this we applied a similar methodology used in the determination of the distance-size ratio $D$.
Using different values of $\delta$, ranging from 0.1 to 1 with a step of 0.1, we computed the potential, the $g_z$ component of the gradient and the $g_{zz}$ component of the Marussi tensor generated by a spherical shell with several exponential densities (like the one in equation \ref{eq:density-exp}): each one with a different value of $b$.
We performed this tests for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.

Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a show the exponential density functions used on each test for a thin and thick spherical shell, respectively.
While on Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b we can see the differences between the resulting $g_{zz}$ with the corresponding analytical solution.
The results for the other gravity fields were excluded from the figures due to their similarities with the ones for $g_{zz}$.

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thin.pdf}
\caption{
    Results of the delta ratio determination, applied to a thin spherical shell (1km of thickness) with various exponential density functions, each one with different value of $b$.
    (a) Density functions used to perform the test.
    (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
\label{fig:exponential-delta-thin}
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.9\linewidth]{figures/exponential-delta-thick.pdf}
\caption{
    Results of the delta ratio determination, applied to a thick spherical shell (35km of thickness) with various exponential density functions, each one with different value of $b$.
    (a) Density functions used to perform the test.
    (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
\label{fig:exponential-delta-thick}
\end{figure}

In the thin shell case we can conclude that a delta equal to 0.3 guarantees an acceptable accuracy under 0.1\% for the entire set of the density functions, while the thick shell case needs at least a delta equal to 0.2.
We will keep the more conservative value of 0.2 as the default value of delta on future computations.

Furthermore, we computed the difference between the numerical model and the analytical solution in case of a discontinuous density (as the one in equation \ref{eq:density-discontinuous} with $R_c = (R_1 + R_2)/3$) using a $\delta = 0.2$.
The result for the $g_{zz}$ component of the Marussi tensor is bellow the 0.1\% threshold, achieving a good performance for this default value of $\delta$.

>>>>

%~ \subsubsection{Linear Density}

%~ If the density of the spherical shell is a linear function on the radial coordinate, i.e.:

%~ \begin{equation}
    %~ \rho(r') = ar' + b,
%~ \end{equation}

%~ \noindent then the gravity potential generated on any computation point at distance $r$ from its centre can be obtained from the following expression:

%~ \begin{equation}
    %~ V_\text{sh}^\text{lin}(r) = \pi G \left[ 
    %~ a \frac{R_2^4 - R_1^4}{r} +
    %~ b \,\frac{4}{3} \frac{R_2^3 - R_1^3}{r} \right].
    %~ \label{eq:shell-pot-linear}
%~ \end{equation}

%~ The first term on equation \ref{eq:shell-pot-linear} reproduces the potential generated by a spherical shell with variable density $\rho(r') = ar'$, while the second one constitutes the potential generated by a spherical shell with homogeneous density $\rho = b$ \citep{Mikuska2006,Grombein2013}.
%~ So the potential can be read as the one generated by a combination of both shells.

%~ \subsubsection{Exponential Density}

%~ If, on the other hand, the spherical shell has an exponential density in the radial coordinate,

%~ \begin{equation}
    %~ \rho(r') = Ae^{-(r' - \Delta h)/b},
%~ \end{equation}

%~ \noindent the potential (calculated through SymPy \citep{sympy2017}) can be written as follows:

%~ \iftwocol{
%~ \begin{equation}
    %~ \begin{split}
        %~ V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
        %~ Ab 
        %~ \Big[
        %~ & (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1 - \Delta h}{b}} - \\
        %~ & (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2 - \Delta h}{b}}
        %~ \Big].
    %~ \end{split}
%~ \label{eq:shell-pot-exp}
%~ \end{equation}
%~ }{
%~ \begin{equation}
    %~ V_\text{sh}^\text{exp}(r) = \frac{4\pi G}{r} 
    %~ Ab \, e^\frac{\Delta h}{b}
    %~ \Big[
    %~ (R_1^2 + 2R_1 b + 2b^2)e^{-\frac{R_1}{b}} - \\
    %~ (R_2^2 + 2R_2 b + 2b^2)e^{-\frac{R_2}{b}}
    %~ \Big].
%~ \label{eq:shell-pot-exp}
%~ \end{equation}
%~ }

@@

%~ \section{Determination of distance-size ratio}

%~ In order to determine the minimum value for the distance-size ratio $D$ such the numerical approximation achieves an acceptable accuracy, we have compared the numerical model with the analytical solutions for a spherical shell with variable density for three special cases: a linear, an exponential and a discontinuous density function.
%~ The first one constitutes the simplest variation, the second one represents a higher varying function, while the later one tries to emulate a contact between two different media (e.g. crust and mantle interface).

%~ We approximated the spherical shell by a mesh of $30^\circ \times 30^\circ$ tesseroids and calculated the gravity potential, the vertical component of the gradient ($g_z$) and the diagonal components of the Marussi tensor ($g_{xx}$, $g_{yy}$, $g_{zz}$) on four different grids: (1) a grid located at the pole, (2) another one on the Equator, (3) one located at satellite height and (4) a big grid of $30^\circ \times 30^\circ$.
%~ More information about this grids can be found on Table \ref{tab:grids}.

%~ We computed these gravity fields on every point of the four grids for different values of $D$, ranging from 0.5 to 10 with a step of 0.5, without applying the density-based discretization algorithm in order to isolate the behaviour of the adaptive discretization one.
%~ Then we calculated the maximum difference between these results and the value that the analytical solution assumes at the same height.

%~ Because the gravity fields generated by a tesseroid are approximated by the effect of point masses, the numerical results may vary between computation points at same height but at different latitude or longitude.
%~ That's why we choose for comparison the maximum difference between the results on every point of the grid and the analytical solution.

%~ Finally, we set the acceptable value of $D$ if the corresponding error of the numerical approximation is lower than 0.1\%.

%~ \begin{table}
%~ \caption{
    %~ Description of the synthetic grids on which the comparison of the numerical model against the analytical solutions for a spherical shell was done. This collection of grids tries to include every possible scenario on which the numerical model can present a different behaviour depending on the size of the grid, its location or its height over the Earth surface. Every grid consists in a set of 10$\times$10 points.
%~ }
%~ \label{tab:grids}
%~ \begin{tabular}{lcccc}
    %~ Grid & Size & Lat. extension & Lon. extension & Height \\ \hline
    %~ Pole & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 2km \\
    %~ Equator & $1^\circ \times 1^\circ$ & $0^\circ - 1^\circ$ & $0^\circ - 1^\circ$ & 2km \\
    %~ Satellite & $1^\circ \times 1^\circ$ & $89^\circ - 90^\circ$ & $0^\circ - 1^\circ$ & 260km \\
    %~ Big Grid & $30^\circ \times 30^\circ$ & $60^\circ - 90^\circ$ & $0^\circ - 30^\circ$ & 2km \\
%~ \end{tabular}
%~ \end{table}

%~ \subsection{Linear Density}
%~ Firstly,  we proposed a spherical shell with linear density in such way it assumes 2670~kg/m$^3$ on the surface of the Earth and 3300~kg/m$^3$ at the inner radius:

%~ \begin{equation}
    %~ \rho(r') = ar' + c,
    %~ \label{eq:density-linear}
%~ \end{equation}
%~ \noindent with 
%~ \begin{equation}
    %~ a = -\frac{3300\text{kg/m$^3$} - 2670\text{kg/m$^3$}}{R_2 - R_1},
%~ \end{equation}
%~ \begin{equation}
    %~ c = \frac{3300\text{kg/m$^3$} - 
        %~ 2670\text{kg/m$^3$}}{R_2 - R_1} R + 
        %~ 2670\text{kg/m$^3$},
%~ \end{equation}

%~ \noindent where $R$ = 6378.137 km is the mean Earth radius.

%~ Then we computed the difference between the numerical model with its corresponding analytical solution (equation \ref{eq:density-linear}) for a thick and a thin spherical shell of 1km and 35km of thickness, respectively.

%~ We can see the results for the potential, the vertical component of the gradient $g_z$ and the $g_{zz}$ component of the Marussi tensor on Figures \ref{fig:D-linear-thin} and \ref{fig:D-linear-thick}.
%~ We exclude the results corresponding to the other two diagonal components of the tensor because they are similar to the ones obtained for the $g_{zz}$, although they can be found on the repository. \todo[inline]{chequear}

%~ For the linear density case, we can observe that the calculation of the potential and the $g_z$ component of its gradient needs a value of $D=1$ and $D=2$, respectively, in order to ensure an accuracy of 0.1\%.
%~ On the other hand, the $g_{zz}$ computations show a noticeable difference between the thin and the thick shell: the first one needs a value of $D$ equal to 8 while the later only a $D$ of 3.5.
%~ We will keep the more conservative result of $D$ equal to 8 in case of computing the Marussi tensor components in case of a linear density tesseroid.

%~ \begin{figure}
%~ \centering
%~ \includegraphics[width=0.9\linewidth]{figures/linear-D-thin.pdf}
%~ \caption{
    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the linear density defined on equation \ref{eq:density-linear}. The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
%~ }
%~ \label{fig:D-linear-thin}
%~ \end{figure}

%~ \begin{figure}
%~ \centering
%~ \includegraphics[width=0.9\linewidth]{figures/linear-D-thick.pdf}
%~ \caption{
    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the linear density defined on equation \ref{eq:density-linear}. The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
%~ }
%~ \label{fig:D-linear-thick}
%~ \end{figure}

%~ \subsection{Exponential Density}

%~ To test the model in case of an exponential density we proposed the following function:

%~ \begin{equation}
    %~ \rho(r') = A e^{-(r' - R)/b} + C,
%~ \label{eq:density-exp}
%~ \end{equation}
%~ \noindent where
%~ \begin{equation}
    %~ A =
    %~ (3300 \text{kg/m}^3 - 2670 \text{kg/m}^3)
    %~ \left( e^{( R_2 - R_1 )/b} - 1 \right)^{-1},
%~ \end{equation}
%~ \begin{equation}
    %~ C =
    %~ 2670 \text{kg/m}^3 - A,
%~ \end{equation}

%~ \noindent and the value of $b$ equal to the thickness of the spherical shell we will use to perform the test.
%~ In this way, the density function assumes the values of 2670 kg/m$^3$ and 3300 kg/m$^3$ on the shell's outer and inner radii, respectively.
%~ The analytical solution of the gravity potential is easily obtained as the sum of the ones for a pure exponential function (like the one in equation \ref{eq:density-exp}) and for an homogeneous spherical shell.

%~ We have performed this test also with a thin and a thick shell of 1 km and 35 km, respectively, without applying the density-based discretization algorithm.
%~ On both cases, the constant $b$ is equal to the thickness of the shell: 1km for the thin shell and 35km for the thick one.
%~ The results can be seen in Figures \ref{fig:D-exp-thin} and \ref{fig:D-exp-thick}.
%~ Once again, we exclude the ones corresponding to $g_{xx}$ and $g_{yy}$ due to their similarity with the ones obtained for $g_{zz}$, although they can be found in the repository.

%~ \begin{figure}
%~ \centering
%~ \includegraphics[width=0.9\linewidth]{figures/exponential-D-thin.pdf}
%~ \caption{
    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the exponential density defined on equation \ref{eq:density-exp}. The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
%~ }
%~ \label{fig:D-exp-thin}
%~ \end{figure}

%~ \begin{figure}
%~ \centering
%~ \includegraphics[width=0.9\linewidth]{figures/exponential-D-thick.pdf}
%~ \caption{
    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thick spherical shell of 35km of thickness with the exponential density defined on equation \ref{eq:density-exp}. The computations were performed on the four grids described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$ and without applying the density-based discretization algorithm. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
%~ }
%~ \label{fig:D-exp-thick}
%~ \end{figure}

%~ In the same way we analysed the linear density results, we can summarise that the thin spherical shell presents the more conservative values of the distance-size ratio: $D=1$ for the gravity potential, $D=2$ for its gradient components and $D=8$ for the Marussi tensor components.

%~ Moreover, we can see that the differences for the potential and the $g_z$ component of the gradient keep constant even for high values of $D$.
%~ Our interpretation is that the modified adaptive discretization algorithm is limited when increasing the accuracy for this density function, supporting the need of a density-based discretization algorithm.

%~ Because the constant $b$ controls the variation of the density function, we wanted to test how does the accuracy of the method behaves at different values of $b$ when the density-based discretization algorithm is not applied.
%~ In order to do this we proposed a thin spherical shell (1km of thickness) with a set of density functions, each one with a different value of $b$.
%~ We computed the difference between the Marussi tensor component $g_{zz}$ generated by the numerical model on the big grid of 30$^\circ\times$30$^\circ$ (see Table \ref{tab:grids}) and the corresponding analytical solution.
%~ These differences were calculated for each proposed density function and for the same set of values of $D$ we have been using on the previous tests, obtaining one curve for each value of $b$.

%~ On Figure \ref{fig:D-exp-power-thin}a we can see the density functions used to perform these tests and its corresponding results on \ref{fig:D-exp-power-thin}b .
%~ Taking into account that lower values of $b$ implies a more varying function, we can see that for high varying functions the numerical model does not perform well: for $b=100$m the difference is almost 100\%.

%~ We can conclude that even when the density-based discretization algorithm is not applied, the numerical approximation performs under the acceptable threshold for values of $b$ greater than the thickness of the shell, i.e. for low density variations.
%~ On the other hand, a density-based discretization algorithm is needed in order to guarantee the accuracy of the approximation in case of a high varying density function.

%~ \begin{figure}
%~ \centering
%~ \includegraphics[width=0.9\linewidth]{figures/exponential-b-thin.pdf}
%~ \caption{
    %~ Results of the distance-size ratio determination for different exponential density functions in case of a spherical shell of 1km of thickness.
    %~ (a) Density functions used in the tests, each one with a different value of $b$.
    %~ (b) Differences between the Marussi tensor component $g_{zz}$ generated by the numerical model and the analytical solution for each density function. Each computation was performed on the big grid described in Table \ref{tab:grids}, for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy.
%~ }
%~ \label{fig:D-exp-power-thin}
%~ \end{figure}

%~ \subsection{Discontinuous Density}

%~ As a final case, we performed the same test with a discontinuous density function:

%~ \begin{equation}
    %~ \rho(r') = 
    %~ \begin{cases}
        %~ \rho_A & \text{if} \, R_1 \leq r' < R_c \\
        %~ \rho_B & \text{if} \, R_c \leq r' \leq R_2 \\
    %~ \end{cases}
%~ \label{eq:density-discontinuous}
%~ \end{equation}

%~ \noindent where $\rho_A$ and $\rho_B$ are constant densities, $R_1$ and $R_2$ are the inner and outer radii of the shell, respectively, and $R_c$ is the radius on which the density change takes place.

%~ The analytical solution for this kind of density is a linear combination of two homogeneous shells with densities $\rho_A$ and $\rho_B$:

%~ \begin{equation}
    %~ V(r) = \frac{4}{3} \frac{\pi G}{r}
    %~ \left[ \rho_A (R_c^3 - R_1^3) +
    %~ \rho_B (R_2^3 - R_c^3) \right]
%~ \end{equation}

%~ Lets assume that $R_c = (R_1 + R_2)/3$ and perform the test for a thin shell of 1km of thickness without applying the density-based discretization algorithm.
%~ Its results can be seen in Figure \ref{fig:D-discont-asym}: up to $D=10$ the difference between the numerical model and the analytical solution is greater than 10\%, what leads us to conclude that a density-based discretization algorithm is needed in order to guarantee the accuracy of the model.

%~ \begin{figure}
%~ \centering
%~ \includegraphics[width=0.9\linewidth]{figures/discontinuous-D-asym.pdf}
%~ \caption{
    %~ Differences between the gravity fields generated by the numerical model and the analytical solution for a thin spherical shell of 1km of thickness with the discontinuous density defined on equation \ref{eq:density-discontinuous}. The computations were performed on the four grids described in Table \ref{tab:grids} and for different values of the distance-size ratio $D$. If the difference is less than 0.1\%, we consider that the model has achieved an acceptable accuracy for the corresponding value of $D$.
%~ }
%~ \label{fig:D-discont-asym}
%~ \end{figure}

@@

%~ \section{Determination of delta ratio}

%~ In the light of the previous results, a density-based discretization algorithm is needed in order to achieve an acceptable accuracy in those cases on which the modified adaptive discretization algorithm is not sufficient to take into account the numerical error due to density variations.

%~ Besides the implementation of this new algorithm, we must determine the maximum value of the delta ratio $\delta$ that produces an accuracy bellow a predefined threshold in order to guarantee the good performance of the numerical model.

%~ To do this we applied a similar methodology used in the determination of the distance-size ratio $D$.
%~ Using different values of $\delta$, ranging from 0.1 to 1 with a step of 0.1, we computed the potential, the $g_z$ component of the gradient and the $g_{zz}$ component of the Marussi tensor generated by a spherical shell with several exponential densities (like the one in equation \ref{eq:density-exp}): each one with a different value of $b$.
%~ We performed this tests for a thin and a thick spherical shell of 1km and 35km of thickness, respectively.

%~ Figures \ref{fig:exponential-delta-thin}a and \ref{fig:exponential-delta-thick}a show the exponential density functions used on each test for a thin and thick spherical shell, respectively.
%~ While on Figures \ref{fig:exponential-delta-thin}b and \ref{fig:exponential-delta-thin}b we can see the differences between the resulting $g_{zz}$ with the corresponding analytical solution.
%~ The results for the other gravity fields were excluded from the figures due to their similarities with the ones for $g_{zz}$.

%~ \begin{figure}
%~ \centering
%~ \includegraphics[width=0.9\linewidth]{figures/exponential-delta-thin.pdf}
%~ \caption{
    %~ Results of the delta ratio determination, applied to a thin spherical shell (1km of thickness) with various exponential density functions, each one with different value of $b$.
    %~ (a) Density functions used to perform the test.
    %~ (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
%~ \label{fig:exponential-delta-thin}
%~ \end{figure}

%~ \begin{figure}
%~ \centering
%~ \includegraphics[width=0.9\linewidth]{figures/exponential-delta-thick.pdf}
%~ \caption{
    %~ Results of the delta ratio determination, applied to a thick spherical shell (35km of thickness) with various exponential density functions, each one with different value of $b$.
    %~ (a) Density functions used to perform the test.
    %~ (b) Differences between the $g_{zz}$ component of the Marussi tensor generated by the numerical model and its corresponding analytical solution for several values of $\delta$ and $b$.}
%~ \label{fig:exponential-delta-thick}
%~ \end{figure}

%~ In the thin shell case we can conclude that a delta equal to 0.3 guarantees an acceptable accuracy under 0.1\% for the entire set of the density functions, while the thick shell case needs at least a delta equal to 0.2.
%~ We will keep the more conservative value of 0.2 as the default value of delta on future computations.

%~ Furthermore, we computed the difference between the numerical model and the analytical solution in case of a discontinuous density (as the one in equation \ref{eq:density-discontinuous} with $R_c = (R_1 + R_2)/3$) using a $\delta = 0.2$.
%~ The result for the $g_{zz}$ component of the Marussi tensor is bellow the 0.1\% threshold, achieving a good performance for this default value of $\delta$.
